Quantifying Spatially Resolved Hydration Thermodynamics Using Grid Inhomogeneous Solvation Theory [Article v\versionnumber]

### 2022-12-22 manuscript/manuscript.tex
Add Franzs new figure, code for the new figure and integration. Replace GISTPP with python for the binding pocket. Add some TODOs, small fixes and changes all over the place. Restructuring is now complete, some details are not done though.

\subsubsection{Visualization the GIST results}
Visualizing the GIST quantities is possible through the OpenDX (.dx) files, either provided directly from the GIST calculation or output by the gisttools library. 
\subsection{Applying GIST for the Streptavidin-Biotin complex}
In this section we will apply GIST to the full complex and discuss each step of the process in more detail to outline pitfalls. We will afterwards introduce advanced analysis workflows which allow you to modify and better visualize the resulting data.
\subsubsection{System Preparation and Equilibration}
If you want to skip this section, run 
\begin{lstlisting}[style=bash]
make equilibration-targets
\end{lstlisting}
in the code folder. This assumes that you have Python and Amber set up properly and that \software{pmemd.cuda} is in the PATH.

@@

\subsubsection{Contribution of Hydration to Binding}
In the next step, calculate the $\dgsolv$ contributions around biotin in each system (biotin, streptavidin, and complex).
Note that our integration region does not fully reach into bulk, and some of the boundary regions are close to streptavidin. 
Therefore, our results for each system will depend on the exact position of the integration boundary. 
To avoid inconsistencies, it is important to choose exactly the same integration region for each system.
Although we kept the systems rigid, the center of mass might have been shifted during pressure equilibration.
Therefore, we align the complex structure to streptavidin and use the shifted biotin coordinates to define the integration region.

@@

Compute the energy (\inlinecode{Eall}) and entropy (\inlinecode{dTSsix}) contributions separately.
You will find that the energy disfavors binding, because we do not yet include the interaction energy between biotin and streptavidin.

You can also plot the difference between the complex and monomer contributions against the distance to biotin, using the complex coordinates for the holo structure.
The expected result is shown in Figure~\ref{fig-dg-sums}.
\todo{This figure is missing -vah}
%\begin{figure}
%	\centering
%	\includegraphics[width=1.0\linewidth]{figures/deltaG-difference.png}
%	\caption{TODO caption}\label{fig-dg-sums}
%\end{figure}

@@

\section{Visualization}

\subsection{Visualizing DX files}
A minimal PyMOL input to visualize a dx file is shown here.
For reference, an input structure is loaded from a file named \inlinecode{streptavidin.pdb}.
The oxygen density is visualized at an isolevel of 2 (i.e., twice the reference density).
The expected output is shown in Figure~\ref{fig-streptavidin_gO}.

% The current font makes the O look like a tofu / placeholder
\begin{lstlisting}[style=pymol]
load streptavidin.pdb
as surface
color gray70, streptavidin
load gist-gO.dx, gO
isomesh gO_high, gO, 2
color slate, gO_high
\end{lstlisting}

@@

Here, the SASA was generated from a simulation of the holo protein structure.
When the ligand is placed in the binding pocked, all atoms are within the SASA.

@@

For example, voxels that are part of the binding pocket can be defined using a ligand molecule and a distance criteria. In \software{gistpp}, voxels that are within the distance criteria from any ligand heavy atom is considered as part of the region. The resulting file is a binary file and can be visualized using \software{pymol}. \todo{Not GISTPP}

\begin{lstlisting}[style=bash]
gistpp -op defbp -i gO.dx -i2 ligand.pdb

-opt const 3.5 -o gO_3.5.dx
\end{lstlisting}

Here, the binding pocket includes all voxels within \SI{3.5}{\angstrom} of the heavy atoms of biotin.

>>>>

This tutorial discusses the visualisation of OpenDX files in section \ref{sec:visualisation.}
\subsection{Applying GIST for the Streptavidin-Biotin complex}
In this section we will apply GIST to the full complex and discuss each step of the process in more detail to outline pitfalls. We will afterwards introduce advanced analysis workflows which allow you to modify and better visualize the resulting data.
\subsubsection{System Preparation and Equilibration}
If you want to skip this section, run 
\begin{lstlisting}[style=bash]
make equilibration-targets
\end{lstlisting}
in the code folder. This assumes that you have Python and Amber set up properly and that \software{pmemd.cuda} is in the PATH.

@@

\subsubsection{Contribution of Hydration to Binding}
\label{sec:binding_contributions}
In the next step, calculate the $\dgsolv$ contributions around biotin in each system (biotin, streptavidin, and complex).
Note that our integration region does not fully reach into bulk, and some of the boundary regions are close to streptavidin. 
Therefore, our results for each system will depend on the exact position of the integration boundary. 
To avoid inconsistencies, it is important to choose exactly the same integration region for each system.
Although we kept the systems rigid, the center of mass might have been shifted during pressure equilibration.
Therefore, we align the complex structure to streptavidin and use the shifted biotin coordinates to define the integration region.

@@

Compute the energy (\inlinecode{Eall}) and entropy (\inlinecode{dTSsix}) contributions separately.
You will find that the energy disfavors binding, because we do not yet include the interaction energy between biotin and streptavidin. To make this easier, gisttool includes a function to integrate in a radius around atom centers:
\begin{lstlisting}[style=python]
centers = {biotin: biotin_x, 
	   strept: strept_x, 
	   compl: compl_x}
names = {biotin: 'Biotin', 
	   strept: 'Streptavidin', 
	   compl: 'Complex'}
cols = ['Eall_dens','dTSsix_dens']
rmax = 12
results = { names[gist] : 
	   gist.integrate_around(
	   		cols, rmax, centers[gist]) 
	   for gist in [biotin, strept, compl]}
display( pd.DataFrame( results ))

\end{lstlisting}
You can also plot the difference between the complex and monomer contributions against the distance to biotin, using the complex coordinates for the holo structure:
\begin{lstlisting}[style=python]
fig, (ax1, ax2) = plt.subplots(1, 2, figsize=(8, 4))
cutoff = 12
ax1.plot(bins, np.cumsum(biotin_rdf), label='biotin')
ax1.plot(bins, np.cumsum(strept_rdf), label='streptavidin')
ax1.plot(bins, np.cumsum(compl_rdf), label='complex')
ax1.legend()
ax1.axvline(cutoff, color='k', linestyle='--')
ax1.set_xlabel('distance to biotin [Å]')
ax1.set_ylabel('$\Delta G$ contributions [Å]')
ax1.grid()

difference = compl_rdf - biotin_rdf - strept_rdf
ax2.plot(bins, np.cumsum(difference))
ax2.axvline(cutoff, color='k', linestyle='--')
ax2.set_xlabel('distance to biotin [Å]')
ax2.set_ylabel('$\Delta \Delta G$ [kcal/mol]')
ax2.grid()
plt.show()
\end{lstlisting}
You can also plot the difference between the complex and monomer contributions against the distance to biotin, using the complex coordinates for the holo structure.
The expected result is shown in Figure~\ref{fig-dg-sums}.
\todo{Missing caption}
\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{figures/deltaG-difference.png}
	\caption{TODO caption}\label{fig-dg-sums}
\end{figure}

@@

\section{Visualization}
\label{sec:visualization}
\subsection{Visualizing DX files}
A minimal PyMOL input to visualize a dx file is shown here.
For reference, an input structure is loaded from a file named \inlinecode{streptavidin.pdb}.
The oxygen density is visualized at an isolevel of 2 (i.e., twice the reference density).
The expected output is shown in Figure~\ref{fig-streptavidin_gO}.
\todo{VJ -> include the PyMOL settings somewhere, this is not the default output}
% The current font makes the O look like a tofu / placeholder
\begin{lstlisting}[style=pymol]
load output/streptavidin/gist.pdb, streptavidin
as surface
color gray70, streptavidin
load output/streptavidin/gist-gO.dx, gO
isomesh gO_high, gO, 2
color slate, gO_high
\end{lstlisting}

@@

Here, the SASA was generated from a simulation of the holo protein structure.
When the ligand is placed in the binding pocked, all atoms are within the SASA.
\todo{I doubt this is the holo/complex, should be the apo streptavidin, right? -vah´}

@@

For example, voxels that are part of the binding pocket can be defined using a ligand molecule and a distance criteria. 
%In \software{gistpp}, voxels that are within the distance criteria from any ligand heavy atom is considered as part of the region. The resulting file is a binary file and can be visualized using \software{pymol}. \todo{Not GISTPP}
In section~\ref{sec:binding_contributions} we show how to integrate over regions such as the binding pocket. To visualize the considered region, we can create a OpenDX file where we set the considered voxels to 1 and all others to 0. Here, we show what this would look like for a radius of \SI{3.5}{\angstrom} around the heavy atoms of biotin. 
%\begin{lstlisting}[style=bash]
%gistpp -op defbp -i gO.dx -i2 ligand.pdb 
%
%-opt const 3.5 -o gO_3.5.dx
%\end{lstlisting}
\begin{lstlisting}[style=python]
def select(traj, sel):
	"Slice a Trajectory by selection mask."
	return traj.atom_slice(traj.top.select(sel))
### Load GIST data
compl = gt.gist.load_gist_file(
	'complex/gist/gistoutlong.dat', 
	struct='complex/gist/gist_nowat.pdb')
### Define ligand atoms and find voxels around them
rmax = 3.5
biotin_mask = 'resname BTN and not element H'
btn = select(compl.struct, biotin_mask).xyz[0] * 10.
ind, _, _ = compl.distance_to_spheres(btn, rmax)

### Set voxels within 3.5A to 1, write out dx file
compl.loc[ind, 'BP'] = 1
compl.save_dx('BP', 'binding_pocket.dx')
\end{lstlisting}
